---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gt3x2csv

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/tarensanders/gt3x2csv/workflows/R-CMD-check/badge.svg)](https://github.com/tarensanders/gt3x2csv/actions)
<!-- badges: end -->

The goal of gt3x2csv is to convert .gt3x files into (raw) csv files so that they can be analysed in other packages such as [GGIR](https://cran.r-project.org/package=GGIR). The goals of this package are:

1. To create output that as closely mimics output from ActiLife as possible.
1. To be orders of magnitude faster than using ActiLife for conversion.

This package is a rewrite of a package previously written by Danilo de Paula Santos ([\@danilodpsantos](https://github.com/danilodpsantos)).

## Why use gt3x2csv instead of ActiLife?

gt3x2csv has a number of advantages over converting files in ActiLife. Firstly, it is substantially faster both on a per-file basis, and overall. This is largely thanks to the great work on the [`AGread`](https://github.com/paulhibbing/AGread)) package, which uses C++ to read the activity data quickly. gt3x2csv is further bolsted by being able to process files in parallel something not available on ActiLife.

As an example of how much faster it is, see the table below. 

Method | One File | Five Files | Thirty Files<sup>[1](#myfootnote1)</sup>
-------------   | ---- | ---- | ----
**Actilife<sup>[2](#myfootnote2)</sup>** | 58s | 4min 55s | 29min 20s
**gt3x2csv<sup>[3](#myfootnote3)</sup><sub>(Sequential)</sub>** | 8s | 40s | 4mins
**gt3x2csv<sub>(Parallel)</sub>** | 9s | 16s | 3mins

<a name="myfootnote1"><sup>1</sup></a> All files were 159MB, or a little over three days of data.<br>
<a name="myfootnote2"><sup>2</sup></a> Using ActiLife v6.11.9 (newer versions might be faster.<br>
<a name="myfootnote3"><sup>3</sup></a> As run on a AMD Ryzen 7 3700X 8-Core CPU; 32GB of RAM.


## What does gt3x2csv do?

gt3x2csv uses [`AGread`](https://github.com/paulhibbing/AGread) to unpack the GT3X file, and formats the output in the same way as ActiLife.

## Installation

gt3x2csv is not (currently) available on CRAN. You can install from GitHub:

``` r
# install.packages("devtools")
devtools::install_github("tarensanders/gt3x2csv")
```

## Example

Using gt3x2csv is simple. You can provide any of a single file to process, a vector of file paths, or a directory. If you do not provide an output directory, the resulting CSV files are stored in the sample place as the originals. Here is an example:

```{r example-setup, include=FALSE, eval=TRUE}
library(gt3x2csv)
temp_dir <- gt3x2csv:::local_dir_with_files()
# my_directory <- file.path(tempdir(), "example dir")
# dir.create(my_directory)
# file.copy(list.files(temp_dir, full.names = TRUE), my_directory, recursive = TRUE)
```


```{r example}
library(gt3x2csv)

my_directory <- gt3x2csv:::local_dir_with_files()

# An example directory with some GT3X files
list.files(my_directory)

gt3x_2_csv(
  gt3x_files = my_directory,
  outdir = NULL, # Save to the same place
  progress = FALSE, # Show a progress bar?
  parallel = TRUE # Process files in parallel?
  )

# Directory now has the new files.
list.files(my_directory)
```

```{r include=FALSE}
unlink(my_directory, recursive = TRUE)
```


# Caveat Emptor

A few warnings for those using gt3x2csv.

### File Sizes

A GT3X file is a zip file containing some metadata (`info.txt`) and a binary file (`log.bin`) with data recorded by the device. The GT3X file is compressed, making it smaller than the raw versions of these files. Converting to CSV uncompresses the file, and will take up more space. This is true regardless of if you use ActiLife or gt3x2csv. How much extra space seems to depend on the size of the file. Here's three files, and their compressed/uncompressed sizes.

File | GT3X Size | CSV Size | Times Larger
---- | ---- | ---- | ----
**File 1** | 200KB | 4.28MB | ~22
**File 2** | 159MB | 524MB | ~3.3
**File 3** | 352MB | 1.13GB | ~3.3

All this is to say that if you can do your analysis without saving CSV files in the middle (e.g., using [`read.gt3x`](https://github.com/THLfi/read.gt3x) or [`AGread`](https://github.com/paulhibbing/AGread)), that would be better. But, some processing packages (e.g., [`GGIR`](https://github.com/wadpac/GGIR)) don't allow this since (for good reasons).

### Memory Use

In the process of unzipping the files, the data are temporarily stored in memory. If you run gt3x2csv in parallel with lots of cores, you might run out of memory. If this happens, just set `cores` to be a lower value.

### Differences to ActiLife

I've validated gt3x2csv against the output from ActiLife using several different files. There are also tests to check that changes to the package do not muck this up. However, this package is provided with no guarantee, and you should test the output yourself.
